### 2017-05-23 sheets/basis-2/sheet-basis-2.tex
Fixed exercise Mostafa (and not me\!) was responsible for.

@@

\begin{answer}
\begin{itemize}
\item We assume mapping a function $f: X \rightarrow Y in$ $\mathbf {Set}$ to a free monoid morphism $F: MonX \rightarrow MonY$ such that $F(Z \in MonX) = (f(z) | z \in Z)$
\item F is faithful since if two sets have the same free monoids (All the lists that can be generated from the elements of those sets), then the two sets are the same.
\item F is not full because we can have a morphism $F(Z \in MonX) = (f(z)f(z) | z \in Z)$
\end{itemize}
\end{answer}

>>>>

\begin{exercise}
  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories. Consider a contravariant functor $F:\A^{\op}\rightarrow \B$. Describe properties of the corresponding function $f:A\rightarrow B$.
\end{exercise}

\begin{answer}
  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
  The functor laws hold tautologically in posets and don't tell us anything about $f$.
\end{answer}

@@

\begin{answer}

We define $F\from\Set\to\Mon$ in the natural way:

\begin{align*}
  F(A)&=A^*\\
  F(f)&=f^*
\end{align*}
where
\begin{align*}
  f^*(a_1\cdots a_n)=f(a_1)\cdots f(a_n).
\end{align*}
Clearly, this is a functor.

Now, suppose $A,A'\in\Set$ and $f,f'\from A\to A'$ such that $F(f)=F(f')$. In particular, for all $a\in A$
\begin{align*}
  f(a)=F(f)(a)=F(f')(a)=f'(a)
\end{align*}
Hence, $f=f'$ and $F$ is faithful.

Next, suppose $A\in\Set$. Define $g\from F(A)\to F(A)$ by $g(a_1\cdots a_n)=a_1a_1\cdots a_na_n$ (i.e. each ``character'' is duplicated). $g$ is a monoid monomorphism but $F(f)\neq g$ for all $f\from A\to A$ since such $F(f)$ in particular preserve lengths of words.

Ergo, $F$ is not full.
\end{answer}
